Stack Safety as a Security Property

### 2020-06-29 main.tex
Quick read of first two sections, couple of notes

But what {\em is} stack safety? A precise formal definition of stack
safety is lacking from the literature, with the notion only being
defined negatively: through examples of stack unsafety. Two classical
examples of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
\leo{Goal: on the left, a bad read; on the right, a bad write.}
\leo{Do we also need an example of ROP/control flow for the new part?}
Recently, a \leo{inspiring/awesome/something good} work reasoned
about a stack safety enforcement mechanism using capabilities\leo{cite};
while impressive, stack safety was defined with respect to an abstract
machine that ``exhibits stack safety by construction''. Which begs
the question, what {\em is} the property that holds by construction?

\leo{StackTokens} provides some intuition towards answering that question,
by informally describing two components that comprise stack safety:
local state encapsulation and well-bracketed control flow. In this paper
we turn this intuition into two formal \leo{extensional} security properties,
that together form a formal definition of stack safety.

@@

While integrity and confidentiality are relatively standard security
properties, this setting provides an interesting additional challenge:
the notion of who the attacker shifts throught execution. Not only
does a caller need to be protected from its callee, but the callee
itself needs to be potentially proteced from {\em its} callees.  In
this paper we address this challenge by varying the observation power
of the attacker at call points, and demonstrate that our policy is
implementable by pre-existing monitoring mechanisms.

@@

Before we define stack safety, we need to first set the stage by
describing our machine model. Concretely, we have in mind a
conventional machine architecture with a standard (RISC-V) processor
extended with some kind of (static or dynamic, software- or
hardware-enforced) security monitor. However, our definition of stack
safety is generic: we will begin with an abstract interface to our
machine model and only make specific choices about details when we get
to the policy enforcement section~\ref{sec:micropolicy}.

@@

The basic building blocks of the machine are {\em values} and {\em
  addresses}. Both are drawn from some set of all possible words
$\WORDS$. We will denote words with $\word$.
%
Our machine states are comprised of {\em components}, that can be word
(denoting memory addresses) or registers (drawn from some set $\REGS$.
%
    \[\reg \in \REGS\]
    \[\component \in \COMPONENTS = \WORDS + \REGS \]
%
The registers must include two special purpose registers: the program
counter \PCname and the stack pointer \SP.
%
Formally, a machine state is just a map from components to values:
%
\[\mach \in \MACHS = \COMPONENTS \rightarrow \WORDS\]

@@

A trace over some type $A$ is a potentially infinite, nonempty
sequence of $A$s. We will mostly be concerned with sequences of
machines states, usually denoted $\machT$, and with sequences of
triples of (machine states, policy states, and observations), usually
denoted $\MPT$. To continue, we will require three generic operations
on traces: a ``trace-of'' operation to construct the potentially
infinite traced induced by a step relation; a ``split'', to break a
trace into two components based on a predicate; a ``join'', the
reverse of split; and a ``prefix'', as a shorthand to only keep the
first component of a split.

@@

Let $\callmap$ be a call map and $\codemap$ a code map. We say that a
system satisfies {\em stack confidentiality} with respect to
$\callmap$ and $\codemap$ if, for every initial machine-policy state
pair $(\mach_0, \pol_0)$ and its induced trace $MP$, and every
$n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ to a callee
$f$ in $MP$, the prefix $MP_{pre}$ of $MP$ from $(\mach_e, \pol_e)$
until the first return corresponding to $\mach_c$ (if any) satisfies
trace integrity with respect to every similarly prefixed machine trace
of every variation $\mach_e'$ with the contour at the call:
$$\forall MP_{pre} ~ M_{pre}. ~
\mpstate{\mach_e}{\pol_e} \hookrightarrow_P MP_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
$$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, f, n)} \mach_e' \wedge
\mach_e' \hookrightarrow M_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ MP_{pre} ~ M_{pre}$$

>>>>

But what {\em is} stack safety? A precise formal definition of stack
safety is lacking from the literature, with the notion only being
defined negatively: through examples of stack unsafety
\rb{usually though not always; as noted below there are other approaches}. Two classical
examples of stack unsafety appear in Figure~\ref{fig:stackunsafety}.
\leo{Goal: on the left, a bad read; on the right, a bad write.}
\leo{Do we also need an example of ROP/control flow for the new part?}
Recently, a \leo{inspiring/awesome/something good} work reasoned
about a stack safety enforcement mechanism using capabilities\leo{cite};
while impressive, stack safety was defined with respect to an abstract
machine that ``exhibits stack safety by construction''. Which begs
the question, what {\em is} the property that holds by construction?

\leo{StackTokens} provides some intuition towards answering that question,
by informally describing two components that comprise stack safety:
local state encapsulation and well-bracketed control flow. In this paper
we turn this intuition into two formal \leo{extensional} security properties,
that together form a formal definition of stack safety.
\rb{I think the merit of StkTokens goes beyond mere intuition. There is a lot of
rigor in it, even if some notions are implicit or not fully developed.}

@@

While integrity and confidentiality are relatively standard security
properties, this setting provides an interesting additional challenge:
the notion of who the attacker is shifts throught execution. Not only
does a caller need to be protected from its callee, but the callee
itself needs to be potentially proteced from {\em its} callees.  In
this paper we address this challenge by varying the observation power
of the attacker at call points, and demonstrate that our policy is
implementable by pre-existing monitoring mechanisms.

@@

Before we define stack safety, we need to first set the stage by
describing our machine model. Concretely, we have in mind a
conventional machine architecture with a standard (RISC-V) processor
extended with some kind of (static or dynamic, software- or
hardware-enforced) security monitor
\rb{Put stress on genericity first, advance this concrete model after?}.
However, our definition of stack
safety is generic: we will begin with an abstract interface to our
machine model and only make specific choices about details when we get
to the policy enforcement section~\ref{sec:micropolicy}.

@@

The basic building blocks of the machine are {\em values} and {\em
  addresses}. Both are drawn from some set of all possible words
$\WORDS$. We will denote words with $\word$.
%
Our machine states are comprised of {\em components}, that can be word
(denoting memory addresses) or registers (drawn from some set $\REGS$).
%
    \[\reg \in \REGS\]
    \[\component \in \COMPONENTS = \WORDS + \REGS \]
%
The registers must include two special purpose registers: the program
counter {\PCname} and the stack pointer \SP.
%
Formally, a machine state is just a map from components to values:
%
\[\mach \in \MACHS = \COMPONENTS \rightarrow \WORDS\]

@@

A trace over some type $A$ is a potentially infinite, nonempty
sequence of $A$s. We will mostly be concerned with sequences of
machines states, usually denoted $\machT$, and with sequences of
triples of (machine states, policy states, and observations), usually
denoted $\MPT$.
\rb{The way the definitions are set up, one may expect MPO, but\ldots}
To continue, we will require three \{four?} generic operations
on traces: a ``trace-of'' operation to construct the potentially
infinite traced induced by a step relation
\rb{not an op on traces per se?};
a ``split'', to break a
trace into two components based on a predicate; a ``join'', the
reverse of split; and a ``prefix'', as a shorthand to only keep the
first component of a split.

@@

Let $\callmap$ be a call map and $\codemap$ a code map.
\rb{Could we find better symbols for the maps?}
We say that a
system satisfies {\em stack confidentiality} with respect to
$\callmap$ and $\codemap$ if, for every initial machine-policy state
pair $(\mach_0, \pol_0)$ and its induced trace $MP$, and every
$n$-transition $(\mach_c, \pol_c)$ and $(\mach_e, \pol_e)$ to a callee
$f$ in $MP$, the prefix $MP_{pre}$ of $MP$ from $(\mach_e, \pol_e)$
until the first return corresponding to $\mach_c$ (if any) satisfies
trace integrity with respect to every similarly prefixed machine trace
of every variation $\mach_e'$ with the contour at the call:
$$\forall MP_{pre} ~ M_{pre}. ~
\mpstate{\mach_e}{\pol_e} \hookrightarrow_P MP_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
$$\forall \mach_e'. ~ \mach_e \approx_{\mathit{Cof}(\mach_c, f, n)} \mach_e' \wedge
\mach_e' \hookrightarrow M_{pre} ~ | ~ \ret{\mach_c} \rightarrow$$
$$\mathit{conf}~\mathit{Cof}(\mach_c, f, n) ~ MP_{pre} ~ M_{pre}$$
